DPIV Analysis and Real Time Implementation

### 2023-01-09 Paper/main.tex
adding comment to fig. 2
Add some changes

\section{Materials and Methods}
This chapter of the paper is organized as follow. \sect\ref{sec:design} provides a schematic description of the hardware platform realization, \sect\ref{sec:method} describes the applied methodology and in \sect\ref{sec:setup} the experimental setup and campaigns are presented.

@@

From a methodological point of view, the experimental characterization of the process takes place through the acquisition of videos that allow studying different interaction mechanisms between particles/cells subjected to different external stresses.
The captured videos are then analyzed using Digital Particle Image Velocimetry (DPIV).
In order to study the different interaction mechanisms between particles in micro-channels subject to external hydrodynamic actuation, the system is composed of i) the hydrodynamic actuation system, ii) the optomechanical system, iii) the microfluidic device,, iv) a video-acquisition system, v)  a PC where the algorithm runs, as shown in \fig\ref{Platform}.
\DS{add the task for each element, i.e. the hydrodynamic system to move the fluids, the optomechanical system to produce the desired magnification, and so on}

@@

\DS{OK ---- Provide a little bit of scientific context where DPIV has been used. I would try to make a small search if a part from this work~\cite{2017-CaiSanBucOrtCabInt}, offline/online versions have been presented.  }
In this paper, a DPIV online platform implementation for data acquisition and processing is proposed. Starting from the DPIV  offline implementation proposed in.....\DS{OK ---- add reference and refer to the previous work}, a the online implementation to ... \DS{add the benefits of the online algorithm}. The key distinction between the two approaches is that the offline algorithm operates on a video captured at an earlier time and uses a slight different software platform than the one employed for the analysis.
~\fig\ref{Algorithms} shows a schematic of the two implementations and highlights the main differences.

@@

Both implementations share the initial steps of establishing the camera's settings (\fig\ref{Algorithms} block (1) \DS{no need to write "block (1)" just "(1)" is OK}) by making adjustments to the pixel size parameters listed in the camera datasheet, the magnification, the exposure time and the image gain. In particular, the latter is an amplification factor applied to all pixel values used for adjusting the pictures brightness level.
The online version differs significantly from the offline one due to the fact that the acquisition takes place in real time through a live stream from the CCD camera to the personal computer.  (\fig\ref{Algorithms} block (4)). This allows to capture process frames directly from the MATLAB routine and they are ready to be used for the analysis (\fig\ref{Algorithms} block (5)){We said to be quite general at this stage. So MATLAB for the moment shouldn't be mentioned}. The offline version, on the other hand, provides for a further step of uploading on MATLAB the video previously acquired through a dedicated software (\fig\ref{Algorithms} block (2)) and unpack it into the corresponding frames (\fig\ref{Algorithms} block (3)). 
Online frames acquisition through MATLAB allows to obtain a higher resolution process frame for the same camera parameters as it is shown in ~\fig\ref{frame}\DS{OK ---- to be verified}.

@@

An on-screen Region of Interest (ROI) selection procedure (\fig\ref{Algorithms} block (6)) allows setting the left-top corner of the image as the starting pixel and the right-down corner of the image as the end pixel. The resulting coordinates for the selection of the ROI are represented in ~\fig\ref{frame}: the abscissa ($\abscissa$) and ordinate ($\ordinate$) of the left-top point and the width ($\width$) and height ($\height$) of the region itself, obtained from the selection of the right-down point.\DS{The ROI choice is important but not at the point we use 9 rows to descibe its choice.} 

@@

The first analysis concerns a DPIV-based algorithm (\fig\ref{Algorithms} block (8)) that provides instantaneous velocity measurements and visualization. Briefly summarizing the DPIV approach presented in ... \MANU{Add reference},it consists of a first collection of the time-varying velocity vector maps through the evaluation of the cross-correlation between ROIs of consecutive pairs of images. 

@@

The preliminary step of DPIV setting (\fig\ref{Algorithms} block (7)) concerns the possibility of opting for a multi-pass discrete Fourier transform (DFT) for the DPIV analysis, by setting the number of passes from 1 to 4, the corresponding interrogation areas and the step size.
A multi-pass DFT in frequency domain, as implemtend by the PIVlab tool\DS{Again, let's keep the generality of the discussion}, could be used in order to increase the accuracy. In the experimental campaign, detailed in \sect\ref{sec:setup}, the analysis was conducted by a three-passes DFT, reaching a good compromise between the resolution and the computational time. The three interrogation areas in pixels were chosen as follows: $Area1=64$, $Area2=32$ and $Area3=16$. The step size was set equal to the half of the last interrogation area ($Area3=16$), resulting in a value equal to 8 $\pixel$.

@@

The micro-particles hydrodynamic response in frequency domain was evaluated by computing the spectrum of $\meann$ in the two components $\abscissa$ and $\ordinate$. This provides informations on the frequency components involved in the dynamic process in terms of fundamental frequency, imposed by the hydrodynamic input and to investigate possible high-frequency components to be associated with intrinsic micro-particles behaviors\DS{I wouldn't use fundamental frequency. I'd rather stress the fact that it comes from the input signal }.
\\The second analysis that this platform provides is the counting particles one (\fig\ref{Algorithms} block (9)). The procedure is based on highlighting the particles in the image distinguishing them from the background. It provides a continuous counting in the time of the micro-particles number in the investigated area, avoiding any manual and individual selecting of the frames to be studied \DS{Although current approaches rely on manual counting or others, (add references if any) the provided implementation makes use of a counting algorithm ...}. It allows to analyze the video frame by frame, count the number of particles and collect this information in a signal. This last\DS{The latter} can also be averaged over time to obtain a single average particles number for a given experimental condition.
\\The counting procedure is essential not only to have an idea of the number of micro-particles involved in the process, but, even more, when we want to compare the results obtained with different types of particles\DS{The structure "when we want" is not that formal. Try to rewrite this sentence}. The velocity obtained by the DPIV-based algorithm, in fact, can be strongly influenced by the number of particles present in the ROI since a spatial average is considered. It should also be considered that different micro-particles correspond to different densities which influence the number of them in the process. According to what has been said, it is necessary to normalize the velocity signals for the average particles number when we wants to compare processes characterized by micro-particles of different nature and physical properties\DS{Try to be more formal and try to shrink a little bit this period about the normalization. }.

@@

A CCD camera (340M Fast Frame, Thorlabs, Newton, NJ, USA) with a resolution of $640 \times 480$ \pixel (pixel size of 7.4~$[\rm\mu m]$, square) was included in the optomechanical system ((OTKB/M) Modular Optical Tweezers, Thorlabs, Newton, NJ, USA). Visible light from the LED source illuminates the sample and is then imaged on the CCD camera.
The CCD camera was connected through a USB connection with a PC for the data acquisition and the subsequent analysis phase in the dedicated software platform.

\begin{figure*}[t]
	\centering
	\includegraphics[width=2\columnwidth]{images/setup}
	\centering{\caption{\label{setup}A picture showing the experimental set-up.}}
\end{figure*}

@@

\section{Results and discussion}
In this chapter\DS{section} the experimental results are presented. In particular in \sect\ref{sec:suspended-particles} we will discuss the synthetic particles behavior related to the fluid in which they are suspended and in \sect\ref{sec:comparison} we will compare the hydrodynamic response of the different types of cells, viable, dead, and synthetic particles. \sect\ref{sec:OnlineOffline} shows a comparison of the results obtained with the two implementations presented in \sect{\ref{sec:method}}, i.e. online and offline implementations.
Since micro-particles are mostly induced to move horizontally due to the input flow, this velocity component will be analyzed in the following analysis.
%Looking at the movement of the micro-particles along the micro-channel and considering also the image orientation, the dominant velocity component is the horizonatal one. In fact, the trend of the horizontal velocity component $<V_x(t)>$ shows the periodicity of the oscillating input flow imposed. This component is the one taken in consideration for the following analysis.

@@

The described algorithm \DS{I'd stress the real-time properties of the algorithm. Therefore, "the real-time capabilities of the algorihtm have been used to"} was used to study how the hydrodynamic response of synthetic micro particles, i.e. silica beads, can be related to their physical properties and differ according to the fluid in which they are suspended. To compare the results obtained for the same kind of particles suspended in different fluids, the velocity signal $\mean{x}$ was computed per each experiment, by the DPIV-based algorithm described in \sect\ref{sec:method}.

@@

In ~\fig\ref{Beads}, for silica beads in PBS and in glycerol-water solution, in the experimental condition with $\amplitude=0.1~\flow$ and  $\frequency=0.1~\hertz$, the velocity signals in the dominant component  $<V_x(t)>$ and their spectrum are plotted.
It can be noted that the speed signal relating to the silica beads sample in the PBS solution (see ~\fig\ref{Beads}(a)) is noisier than the speed signal relating to the silica beads sample in the glycerol-water solution (see ~\fig\ref{Beads}(c)). This is given by the fact that since the silica beads have a higher density than PBS, they tend to settle in the bottom of the micro channel, putting up resistance to the imposed hydrodynamic actuation at the input\DS{I wouldn't say that the signal is noisier than.., it may seem that our setup is the source of the problem. The second argument is perfect.}. Suspended beads in a solution of the same density have shown to be an effective solution to this problem. As a result, the hydrodynamic input exerted on the fluid has no effect on the beads' ability to float freely and the speed signal relating to the silica beads sample in the glycerol-water solution (see ~\fig\ref{Beads}(c)) shows a shape similar to the input flow supplied, i.e. a sine wave with frequency $\frequency=0.1~\hertz$. In addition, such frequency can be detected from the spectrumv\DS{add ref picture}, and amplitude $\amplitude=0.1~\flow$.
\\The micro-particles hydrodynamic response in time domain was evaluated by computing the range of the velocity signal $\mean{x}$, as shown in \eqn (\ref{eqn:range}). ~\fig\ref{BeadsFinal} shows the value of $\range$ and the amplitude peak ($\peak$) at the fundamental frequency for each experimental condition. The two curves are related to silica beads in the two different solutions considered and each point of the curves is correlated to the amplitude of the external oscillating input flow strength $\amplitude$ with a constant $\frequency=0.1~\hertz$. 
\\The curve related to the beads in PBS does not show any correlation between the imposed input and the hydrodynamic response of the micro-particles. The curve relating to the beads in the glycerol-water solution shows how the parameters examined in the time domain and in the frequency domain increase as the input amplitude increases, showing a correlation between the input and the hydrodynamic response of the micro-particles.
Since silica beads are synthetic particles, what we expect is that they do not provide any resistance to the input sources and so the glycerol solution is the most suitable among the analyzed ones because it allows to show up this behavior.
For this reason, in \sect\ref{sec:comparison} in which the comparison between synthetic particles, live cells and cells exposed to death stimuli is made, and in \sect\ref{sec:OnlineOffline} in which the results obtained from the online and offline platform are compared, only the results relating to silica beads in the glycerol-water solution will be considered because they allowed a better process characterization\DS{This is an intensely literal translation of an Italian sentence. The difference might be as simple as switching the positions of the subject and verb. Besides, active forms are more effective than passive ones.}.

@@

The DPIV-based and micro-particles counting algorithms were used to analyze the data collected in the experimental campaigns related to yeast cells in different living conditions. This allowed to investigate how the hydrodynamic response of cells could be correlated to their living condition.  To compare the results obtained for different kind of cells' living condition, the velocity signal $\mean{x}$ was computed by the DPIV-based algorithm described in \sect\ref{sec:method}, the spectrum of each signal was evaluated and the average number of micro-particles $\particles$ was computed for each experiment. 
In particular, the results relating to viable yeast cells and viable yeast cells exposed to cell death stimuli, induced by an apoptotic process of death, i.e. a mechanism of genetically programmed death, are reported \DS{TODO: provide additional details in the introduction about apoptosis and add references}. These two conditions allowed to correlate the hydrodynamic response of the micro-particles to their physical properties according to the different living conditions. This allowed a characterization of the same type of cells, the yeast cells, in the different phases of the cell cycle.

@@

In ~\fig\ref{Yeast}, for viable and exposed to cell death stimuli yeast cells\DS{it may be more effective to use a more concise expression for death cells or an acromym that we can define at the beginning. What's your opinion?  }, in the experimental condition with $\amplitude=0.1~\flow$ and  $\frequency=0.1~\hertz$, the velocity signals in the dominant component  $\mean{x}$ are plotted. The different oscillating trends of the velocity signals highlight the effects induced by the oscillating input flow in different cell living conditions. In particular, considering the same experimental condition, the amplitude of the velocity trend for viable yeast cells (see ~\fig\ref{Yeast}(a)) is lower than the one found for yeast cells exposed to cell death stimuli. 
In the condition of viable yeast cells there is a greater resistance from the cells to the force imposed on the input, resulting in a lower maximum speed value reached compared to that of yeast cells exposed to cell death stimuli.

>>>>

\section{Materials and Methods}
This section of the paper is organized as follow. \sect\ref{sec:design} provides a schematic description of the hardware platform realization, \sect\ref{sec:method} describes the applied methodology and in \sect\ref{sec:setup} the experimental setup and campaigns are presented.

@@

From a methodological point of view, the experimental characterization of the process takes place through the acquisition of videos that allow studying different interaction mechanisms between particles/cells subjected to different external stresses.
The captured videos are then analyzed using Digital Particle Image Velocimetry (DPIV).
In order to study the different interaction mechanisms between particles in micro-channels subject to external hydrodynamic actuation, the system is composed of i) the hydrodynamic actuation system to inject the fluid sample inside the micro-channel, ii) the optomechanical system that includes the light sources, the objective lense and all the hardware components used to hold and align elements, iii) the microfluidic device, set of the micro-channel, iv) a video-acquisition system to capture video of the process v)  a PC where the algorithm runs, as shown in \fig\ref{Platform}.

@@

\MANU{ Provide a little bit of scientific context where DPIV has been used. I would try to make a small search if a part from this work~\cite{2017-CaiSanBucOrtCabInt}, offline/online versions have been presented.  }
In this paper, a DPIV online platform implementation for data acquisition and processing is proposed. Starting from the DPIV  offline implementation proposed in.....\MANU{Add reference and refer to the previous work}, an online implementation to compact the phases of capturing and unpacking the video is proposed. The key distinction between the two approaches is that the offline algorithm operates on a video captured at an earlier time and uses a slight different software platform than the one employed for the analysis.
~\fig\ref{Algorithms} shows a schematic of the two implementations and highlights the main differences.

@@

Both implementations share the initial steps of establishing the camera's settings (\fig\ref{Algorithms} (1)) by making adjustments to the pixel size parameters listed in the camera datasheet, the magnification, the exposure time and the image gain. In particular, the latter is an amplification factor applied to all pixel values used for adjusting the pictures brightness level.
The online version differs significantly from the offline one due to the fact that the acquisition takes place in real time through a live stream from the CCD camera to the personal computer.  (\fig\ref{Algorithms} (4)). This allows to capture process frames directly from the routine and they are ready to be used for the analysis (\fig\ref{Algorithms} (5)). The offline version, on the other hand, provides for a further step of uploading on the routine the video previously acquired through a dedicated software (\fig\ref{Algorithms} (2)) and unpack it into the corresponding frames (\fig\ref{Algorithms} (3)). 
Online frames acquisition allows to obtain a higher resolution process frame for the same camera parameters as it is shown in ~\fig\ref{frame}\MANU{To be verified}.

@@

An on-screen Region of Interest (ROI) selection procedure (\fig\ref{Algorithms} (6)) allows setting the left-top corner and the right-down corner of the image as the starting and the ending $\pixel$. The resulting coordinates for the ROI selection are reported in ~\fig\ref{frame}: the abscissa ($\abscissa$) and ordinate ($\ordinate$) of the left-top point and the width ($\width$) and height ($\height$) of the region itself.

@@

The first analysis concerns a DPIV-based algorithm (\fig\ref{Algorithms} (8)) that provides instantaneous velocity measurements and visualization. Briefly summarizing the DPIV approach presented in ... \MANU{Add reference},it consists of a first collection of the time-varying velocity vector maps through the evaluation of the cross-correlation between ROIs of consecutive pairs of images. 

@@

The preliminary step of DPIV setting (\fig\ref{Algorithms} (7)) concerns the possibility of opting for a multi-pass discrete Fourier transform (DFT) for the DPIV analysis, by setting the number of passes from 1 to 4, the corresponding interrogation areas and the step size.
A multi-pass DFT in frequency domain could be used in order to increase the accuracy. In the experimental campaign, detailed in \sect\ref{sec:setup}, the analysis was conducted by a three-passes DFT, reaching a good compromise between the resolution and the computational time. The three interrogation areas in $\pixel$ were chosen as follows: $Area1=64$, $Area2=32$ and $Area3=16$. The step size was set equal to the half of the last interrogation area ($Area3=16$), resulting in a value equal to 8 $\pixel$.

@@

The micro-particles hydrodynamic response in frequency domain was evaluated by computing the spectrum of $\meann$ in the two components $\abscissa$ and $\ordinate$. This provides informations on the frequency components involved in the dynamic process in terms of the one imposed by the hydrodynamic input and to investigate possible high-frequency components to be associated with intrinsic micro-particles behaviors.
\\The second analysis that this platform provides is the counting particles one (\fig\ref{Algorithms} (9)). The procedure is based on highlighting the particles in the image distinguishing them from the background. It provides a continuous counting in the time of the micro-particles number in the investigated area, avoiding any manual and individual selecting of the frames to be studied \MANU{Although current approaches rely on manual counting or others, (add references if any) the provided implementation makes use of a counting algorithm ...}. It allows to analyze the video frame by frame, count the number of particles and collect this information in a signal. The latter can also be averaged over time to obtain a single average particles number for a given experimental condition.
\\The counting procedure is essential not only to have an idea of the number of micro-particles involved in the process, but, even more, to compare the results obtained with particles of different nature. The velocity obtained by the DPIV-based algorithm, in fact, can be strongly influenced by the number of particles present in the ROI since a spatial average is considered. It should also be considered that different micro-particles correspond to different densities which influence the number of them in the process. According to what has been said, it is necessary to normalize the velocity signals for the average particles number when we wants to compare processes characterized by micro-particles of different nature and physical properties\DS{Try to be more formal and try to shrink a little bit this period about the normalization. }.

@@

A CCD camera (340M Fast Frame, Thorlabs, Newton, NJ, USA) with a resolution of $640 \times 480$ $\pixel$ (pixel size of 7.4~$[\rm\mu m]$, square) was included in the optomechanical system ((OTKB/M) Modular Optical Tweezers, Thorlabs, Newton, NJ, USA). Visible light from the LED source illuminates the sample and is then imaged on the CCD camera.
The CCD camera was connected through a USB connection with a PC for the data acquisition and the subsequent analysis phase in the dedicated software platform.

\begin{figure*}[t]
	\centering
	\includegraphics[width=2\columnwidth]{images/setup}
	\centering{\caption{\label{setup}A picture showing the experimental set-up. \MANU{Figure need to be redone} }}
\end{figure*}

@@

\section{Results and discussion}
In this section the experimental results are presented. In particular in \sect\ref{sec:suspended-particles} we will discuss the synthetic particles behavior related to the fluid in which they are suspended and in \sect\ref{sec:comparison} we will compare the hydrodynamic response of the different types of cells, viable, dead, and synthetic particles. \sect\ref{sec:OnlineOffline} shows a comparison of the results obtained with the two implementations presented in \sect{\ref{sec:method}}, i.e. online and offline implementations.
Since micro-particles are mostly induced to move horizontally due to the input flow, this velocity component will be analyzed in the following analysis.
%Looking at the movement of the micro-particles along the micro-channel and considering also the image orientation, the dominant velocity component is the horizonatal one. In fact, the trend of the horizontal velocity component $<V_x(t)>$ shows the periodicity of the oscillating input flow imposed. This component is the one taken in consideration for the following analysis.

@@

The described algorithm \DS{I'd stress the real-time properties of the algorithm. Therefore, "the real-time capabilities of the algorihtm have been used to"} \MANU{Why "real time"?} was used to study how the hydrodynamic response of synthetic micro particles, i.e. silica beads, can be related to their physical properties and differ according to the fluid in which they are suspended. To compare the results obtained for the same kind of particles suspended in different fluids, the velocity signal $\mean{x}$ was computed per each experiment, by the DPIV-based algorithm described in \sect\ref{sec:method}.

@@

In ~\fig\ref{Beads}, for silica beads in PBS and in glycerol-water solution, in the experimental condition with $\amplitude=0.1~\flow$ and  $\frequency=0.1~\hertz$, the velocity signals in the dominant component  $<V_x(t)>$ and their spectrum are plotted.
It can be noted that the speed signal relating to the silica beads sample in the PBS solution (see ~\fig\ref{Beads}(a)) is noisier than the speed signal relating to the silica beads sample in the glycerol-water solution (see ~\fig\ref{Beads}(c)). This is given by the fact that since the silica beads have a higher density than PBS, they tend to settle in the bottom of the micro channel, putting up resistance to the imposed hydrodynamic actuation at the input\DS{I wouldn't say that the signal is noisier than.., it may seem that our setup is the source of the problem. The second argument is perfect.} \MANU{Is it better to use dirtier?}. Suspended beads in a solution of the same density have shown to be an effective solution to this problem. As a result, the hydrodynamic input exerted on the fluid has no effect on the beads' ability to float freely and the speed signal relating to the silica beads sample in the glycerol-water solution (see ~\fig\ref{Beads}(c)) shows a shape similar to the input flow supplied, i.e. a sine wave with frequency $\frequency=0.1~\hertz$ and amplitude $\amplitude=0.1~\flow$. In addition, such frequency can be detected from the spectrum in ~\fig\ref{Beads}(d).
\\The micro-particles hydrodynamic response in time domain was evaluated by computing the range of the velocity signal $\mean{x}$, as shown in \eqn (\ref{eqn:range}). ~\fig\ref{BeadsFinal} shows the value of $\range$ and the amplitude peak ($\peak$) at the fundamental frequency for each experimental condition. The two curves are related to silica beads in the two different solutions considered and each point of the curves is correlated to the amplitude of the external oscillating input flow strength $\amplitude$ with a constant $\frequency=0.1~\hertz$. 
\\The curve related to the beads in PBS does not show any correlation between the imposed input and the hydrodynamic response of the micro-particles. The curve relating to the beads in the glycerol-water solution shows how the parameters examined in the time domain and in the frequency domain increase as the input amplitude increases, showing a correlation between the input and the hydrodynamic response of the micro-particles.
Since silica beads are synthetic particles, what we expect is that they do not provide any resistance to the input sources and so the glycerol solution is the most suitable among the analyzed ones because it allows to show up this behavior.
For this reason, in \sect\ref{sec:comparison}, which shows the comparison between synthetic particles, live cells and cells exposed to death stimuli, and in \sect\ref{sec:OnlineOffline}, which compares the results obtained from the online and offline platform, only the results relating to silica beads in the glycerol-water solution will be considered because they allowed a better process characterization.

@@

The DPIV-based and micro-particles counting algorithms were used to analyze the data collected in the experimental campaigns related to yeast cells in different living conditions. This allowed to investigate how the hydrodynamic response of cells could be correlated to their living condition.  To compare the results obtained for different kind of cells' living condition, the velocity signal $\mean{x}$ was computed by the DPIV-based algorithm described in \sect\ref{sec:method}, the spectrum of each signal was evaluated and the average number of micro-particles $\particles$ was computed for each experiment. 
In particular, the results relating to viable yeast cells and viable yeast cells exposed to cell death stimuli, induced by an apoptotic process of death, i.e. a mechanism of genetically programmed death, are reported \MANU{TODO: provide additional details in the introduction about apoptosis and add references}. These two conditions allowed to correlate the hydrodynamic response of the micro-particles to their physical properties according to the different living conditions. This allowed a characterization of the same type of cells, the yeast cells, in the different phases of the cell cycle.

@@

In ~\fig\ref{Yeast}, for viable and exposed to cell death stimuli yeast cells\DS{it may be more effective to use a more concise expression for death cells or an acromym that we can define at the beginning. What's your opinion?} \MANU{What about simply "dead cells"?}, in the experimental condition with $\amplitude=0.1~\flow$ and  $\frequency=0.1~\hertz$, the velocity signals in the dominant component  $\mean{x}$ are plotted. The different oscillating trends of the velocity signals highlight the effects induced by the oscillating input flow in different cell living conditions. In particular, considering the same experimental condition, the amplitude of the velocity trend for viable yeast cells (see ~\fig\ref{Yeast}(a)) is lower than the one found for yeast cells exposed to cell death stimuli. 
In the condition of viable yeast cells there is a greater resistance from the cells to the force imposed on the input, resulting in a lower maximum speed value reached compared to that of yeast cells exposed to cell death stimuli.
